Freezing of a Weeks-Chandler-Anderson Fluid using Classical Density Functional Theory

### 2021-01-09 papers/thesis-kirstie/thesis.tex
edits

    f(r) = -\theta(|\vec r_i - \vec r_j|-(R_i + R_j))

@@

where and $|\vec r_i - \vec r_j|$ is the distance between the centers of two spheres.

@@

of two hard spheres added together. For a mixture of hard spheres of radius $R_i$ and $R_j$

@@

       \int_{Volume} \theta(|\vec r_i- \vec r_j|-(R_i+R_j))~d\vec r &=\frac{4\pi}{3}(R_i+R_j)^3  \\
       &= \frac{4\pi}{3}R_i^3 + (4\pi R_i^2)R_j + (4\pi R_j^2)R_i + \frac{4\pi}{3}R_j^3  \nonumber \\

@@

        V &= \int_{Volume} \theta(r-R)d\vec r \\

@@

                 &= k_BTn^2\left(\int \frac{\theta(|\vec r|-2R)}{2}~d\vec r\right)V  \\

@@

                 &= k_BT\left[\left(\int n\theta(r-R)d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R^2} d\vec r \right)\color{white} \right]   \nonumber \\
                 &+ \color{white}\left[\color{black}\left(\int n\delta(r-R) d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R} d\vec r \right) \right]\color{black}V   \\          
                 &= k_BT(n_3n_0 + n_2n_1)V \\

@@

                 &= k_BT\Phi(\{n_i\})V

@@

     n_i=\int nw_i(r) d\vec r

@@

      w_3 &= \theta(r-R)  \label{scalor-weight3} \\

@@

and dimensionless energy densities $\Phi_1$ and $\Phi_2$ are defined as
The excess Helmholtz free energy can also be expressed in terms of the
packing fraction $\eta=nV$ as
     F_{excess} &= k_BT(n_3n_0 + n_2n_1)V  \\
                &= k_BT\left(\left(\eta\right)\left(n\right) + \left(\frac{3\eta}{R}\right)\left(\frac{3\eta}{4\pi R^2}\right)\right)V  \\      
                &= k_BT\left(\eta^2+3\eta^2\frac{3}{4\pi R^3}V\right)  \\
                &= k_BT4\eta^2

@@

     F_{excess} = k_BT\int d\vec r_a \int d\vec r_b~n(\vec r_a)~n(\vec r_b)~\frac{\theta(|\vec r_a-\vec r_b|-2R)}{2}

@@

where $\vec r_a$ is a position vector of one sphere, $\vec r_b$ is a position vector to another sphere, and 
$|\vec r_a - \vec r_b|$ is the distance between the centers of the two spheres.

@@

a convolution of the weight functions, 

@@

    f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})

@@

where the weights are the same ones given in Equations~\ref{scalor-weight3} through~\ref{scalor-weight0}, 
and the two vector weights are given by

@@

      \vec {w}_2 &= w_2~\hat r \\
      \vec {w}_1 &= w_1~\hat r

@@

which are needed for the relation to be satisfied (although other expressions 
have been developed without using vector weights). 

@@

He then noted the following relation

@@

     w_\alpha(\vec r_a - \vec r_b) \otimes w_\gamma(\vec r_a - \vec r_b) &= \int w_\alpha(\vec r_a - \vec x) w_\gamma(\vec r_b - \vec x) d\vec x \\
                                                                         &= \int w_\alpha(\vec x - \vec r_a) w_\gamma(\vec x - \vec r_b) d\vec x

@@

which simplifies the equation for the excess Helmholtz free energy

@@

     F_{excess} &= k_BT\int d\vec r_a \int d\vec r_b~n(\vec r_a)~n(\vec r_b)~\frac{\theta(|\vec r_a-\vec r_b|-2R)}{2}  \\
                &= k_BT\int d\vec r_a \int d\vec r_b~n(\vec r_a)~n(\vec r_b)~(w_0\otimes w_3+ w_1 \otimes w_2+\vec {w}_1\otimes \vec {w}_2)  \\
                &= k_BT\int d\vec r_a \int d\vec r_b~n(\vec r_a)~n(\vec r_b)\left(\int w_3(\vec x - \vec r_a)w_0(\vec x - \vec r_b) \vec dx \color{white}\right)  \nonumber \\
                &+ \color{white}\left(\color{black}\int w_1(\vec x - \vec r_a)w_2(\vec x - \vec r_b) d\vec x  + \int \vec w_1(\vec x - \vec r_a)\cdot\vec w_2(\vec x - \vec r_b) d\vec x\right)    \\
                &= k_BT\int \left[\int n(\vec r_a) w_3(\vec x - \vec r_a)~d\vec r_a \int n(\vec r_b) w_0(\vec x - \vec r_b)~d\vec r_b \color{white}\right]  \nonumber \\
                &+ \int n(\vec r_a) w_2(\vec x - \vec r_a)~d\vec r_a \int n(\vec r_b) w_1(\vec x - \vec r_b)~d\vec r_b  \nonumber \\
                &+ \color{white}\left[\color{black} \int n(\vec r_a) \vec {w}_2(\vec x - \vec r_a)~d\vec r_a \cdot \int n(\vec r_b) \vec {w}_1(\vec x - \vec r_b)~d\vec r_b \right]\color{black}d\vec x  \\
                &= k_BT\int n_3n_0 + n_2n_1 + \vec n_1\cdot\vec n_2~d\vec x  \\
                &= k_BT\int \Phi_1 + \Phi_2 + \Phi_3~d\vec x  \\
                & =k_BT\int \Phi(\{n_i(\vec r)\})~d\vec r               

@@

where $\vec n_1\cdot\vec n_2$=0 for the homogeneous case, 
and the weighted densities $n_i$ are defined as

@@

    n_i(\vec r)=\int n(\vec r')w_i(\vec r-\vec r')~d\vec r'

@@

@@

}\textcolor{red}{
Lifting the restriction to the low density limit,
The expression for the excess Helmholtz free energy derived from FMT
is applicable to the more general case of an inhomogeneous fluid, but reduces 
to the same result obtained from Scaled Particle Theory for a homogeneous 
hard-sphere fluid. 

@@

%In order to apply classical Density Functional Theory to a system that has
%a fixed temperature, volume, and number of particles, a functional for 
%the Helmholtz free energy must first be formed. An expression for the ideal 
%portion of the Helmholtz free energy is given by Equation~\ref{eq:Fideal-n(r)}, 
%but forming an expression for the excess Helmholtz free
%energy from Equation~\ref{eq:Fexcess-simplified} for the particular fluid 
%being modeled can be challenging. 
%Soft Fundamental Measure Theory (SFMT) is a method for forming an 
%expression for the excess Helmholtz free energy of a soft-sphere fluid
%that does not depend on explicitly taking into account the interaction 
%that each sphere with every other sphere in the fluid (as required by 
%Equation~\ref{eq:Fexcess-simplified}). In the next chapter on methods, we
%will use an SFMT functional that we adapt to fit the soft-sphere WCA fluid.
%SFMT was developed by Schmidt~\cite{schmidt1999density, schmidt2000fluid} 
%for soft-spheres. It is an extension of Fundamental Measure Theory (FMT) 
%originally developed by 
%Rosenfeld~\cite{rosenfeld1989, rosenfeld1994, rosenfeld1996, fmtfromsptandsignerror} 
%for the hard-sphere fluid, and which expanded 
%Scaled Particle Theory (SPT)~\cite{santos2012phi3} 
%to inhomogeneous, or non-uniform hard-sphere fluids.
%
Scaled Particle Theory
seeks a general expression for the excess chemical 
potential of a homogeneous hard-sphere fluid which can then be integrated 
to find the excess Helmholtz free energy according to the relation
\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~}\rightarrow{~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{equation}
It is an extension of the work done by Widom who showed that the excess 
chemical potential of a homogeneous fluid corresponds to the reversible 
work of inserting a sphere of radius R into a fluid consisting of hard 
spheres of radius R. 
\begin{align}
	\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}
\end{align}
Widom's result was obtained by considering the difference 
in the excess Helmholtz free energy when one atom is added to an otherwise 
closed system. He showed that the excess chemical potential is related to 
the ensemble average of the Boltzman factor for the potential energy 
difference $V_{diff}$ between a system with N atoms and one with N+1 atoms. 
The resulting equation, shown below, is known as Widom's Insertion 
Formula~\cite{Hansen}.
\begin{align} \label{mu-der-int}
    \mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}
\end{align}
\begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
where 
\begin{align}
    <e^{-\beta{V_{diff}}}>=\sum_s{P_se^{-\beta V_s}}
\end{align}
and
\begin{align}
    Z=\sum_s{e^{-\beta V_s}}
\end{align}

@@

Fundamental Measure Theory can be constructed from Equation~\ref{Fexcess-SPT}  
%and provides an expression for the excess Helmholtz free energy of an 
%inhomogeneous 
%hard-sphere fluid as a functional of the number density profile $n(\vec r)$. 
%
%Soft Fundamental Measure Theory (SFMT), which will be used in this paper, 
%is an extension of FMT and provides an expression for the excess Helmholtz
%free energy of fluids with soft potentials like that of the WCA fluid. 
%It will be helpful to continue looking at FMT before addressing SFMT.
%FMT begins 
by noticing that a quantity $G_{3}$ can be defined such that
\begin{equation}{\eta =\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=nG_{3}}\end{equation}
where $G_{3}$ is the volume of a sphere. 
Likewise, by defining quantities
\begin{equation}{G_{2}=4\pi{R^2}}\end{equation}
\begin{equation}{G_{1}=R}\end{equation}
\begin{equation}{G_{0}=1}\end{equation}
where $G_{2}$ is the surface area of a sphere, and $G_{1}$ is the 
radius of a sphere, Equation~\ref{Fexcess-SPT} can be written as
\begin{equation}\label{FexfromSPTinthermsofG}{F_{excess}=k_{B}T\left(-nG_{0}ln(1-nG_{3})+\frac{nG_{1}nG_{2}}{1-nG_{3}}+\frac{nG_{2}^3}{24\pi(1-Gn_{3})^2}\right)V}\end{equation}
The quantities $G_{3}, G_{2}, G_{1},$ and $G_{0}$ are fundamental 
geometric measures of a sphere, and are what give rise to the name
Fundamental Measure Theory. Introducing a set of variables
%weighted densities $n_{i}=nG_{i}$ 
given by
\begin{equation}\label{n3}{n_{3}=nG_{3}=\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=\eta}\end{equation}
\begin{equation}\label{n2}{n_{2}=nG_{2}=\left(\frac{N}{V}\right)4\pi{R^2}=\frac{3\eta}{R}}\end{equation}
\begin{equation}\label{n1}{n_{1}=nG_{1}=\left(\frac{N}{V}\right)R}\end{equation}
\begin{equation}\label{n0}{n_{0}=nG_{0}=\left(\frac{N}{V}\right)}\end{equation}
Equation~\ref{FexfromSPTinthermsofG} can be written as
\begin{align}\label{FexfromSPT}
    F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}}{1-n_{3}}+\frac{n_{2}^3}{24\pi(1-n_{3})^2}\right)V
\end{align}
which has the form 
\begin{align}
    F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V
\end{align}
%where 
\begin{align}
   \Phi_1 &= -n_{0}\ln(1-n_{3}) \\
   \Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
   \Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2}   \label{oldPhi3equation}
\end{align}

The quantities $n_{0},n_{1},n_{2},n_{3}$ can be considered as weighted 
densities where the weights are given by $G_{0},G_{1},G_{2},G_{3}$ respectively.
\begin{align}
    n_{i}=nG_{i}
\end{align}
These weights are valid for determining the excess free energy of a system of 
hard spheres with a homogeneous number density n. But these weights, and 
their related  weighted densities, can be written in a more general form 
allowing for their extension to a system with a spatially varying number 
density $n(\vec{r})$, that is, an inhomogeneous fluid, in which case, the
excess Helmholtz free energy is written as
\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation} 
Using a backwards Heaviside step function defined as
\begin{equation}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{equation}
the following expressions are obtained
\begin{equation}{G_{3}=\int_{Volume}{\Theta(|\vec{r}|-R)d{\vec{r}}} = \frac{4}{3}\pi{R^3}}\end{equation}
\begin{equation}{G_{2}=\int_{Volume}{\delta(|\vec{r}|-R)d{\vec{r}}} = 4\pi{R^2}}\end{equation}
\begin{equation}{G_{1}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r}d{\vec{r}}} = R}\end{equation}
\begin{equation}{G_{0}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r^2}d{\vec{r}}} = 1}\end{equation}
where the weighted densities can now be expressed as
\begin{align}
    n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}
\end{align} 
or for a spatially varying number density $n(\vec{r})$,
\begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
with weight functions
\begin{equation}\label{eq:w3}{w_{3}=\Theta(|\vec{r}|-R)}\end{equation}
\begin{equation}\label{eq:w2}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}
\begin{equation}\label{eq:w1}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
\begin{equation}\label{eq:w0}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
The weight functions $w_{2}$ and $w_{3}$ have the relation
\begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}
which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.

@@

Tarazona later improved upon the original FMT equations by replacing the
expression for $\Phi_3$ given by Equation~\ref{oldPhi3equation} with the
more accurate expression~\cite{tarazonaphi3, santos2012phi3} 
\begin{align}
\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  \label{eq:Phi3}
\end{align}
in which a tensor weight is incorporated given by
\begin{equation}\label{eq:tensorweight}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
All the other FMT equations remain the same.

>>>>

    f(r) = -\Theta(r-(R_i + R_j))

@@

where $r=|\vec r_i - \vec r_j|$ is the distance between the centers of two spheres.

@@

of two hard spheres added together. For a mixture of hard spheres of radius $R_a$ and $R_b$

@@

       \int_{Volume} \Theta(r-(R_i+R_j))~d\vec r &=\frac{4\pi}{3}(R_i+R_j)^3  \label{exclusion-volume}\\
       &= \frac{4\pi}{3}R_i^3 + (4\pi R_i^2)R_j + R_i(4\pi R_j^2) + \frac{4\pi}{3}R_j^3  \nonumber \\

@@

        V &= \int_{Volume} \Theta(r-R)d\vec r \\

@@

                 &= k_BTn^2\left(\int \frac{\Theta(r-2R)}{2}~d\vec r\right)V  \\

@@

                 &= k_BT\left[\left(\int n\Theta(r-R)d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R^2} d\vec r \right)\color{white} \right]   \nonumber \\
                 &+ \color{white}\left[\color{blue}\left(\int n\delta(r-R) d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R} d\vec r \right) \right]\color{blue}V   \\          
                 &= k_BT(n_3n_0 + n_2n_1)V \label{Fexlowden-homo} \\

@@

                 &= k_BT\Phi(\{n_\alpha\})V
The quantity $\Phi$ has dimensions of $V^{-1}$ and is an energy density 
with a dimensionless energy composed 
of the sum of $\Phi_1$ and $\Phi_2$ which are defined as

@@

     n_\alpha=\int nw_\alpha(r) d\vec r

@@

      w_3 &= \Theta(r-R)  \label{scalor-weight3} \\

@@

%The excess Helmholtz free energy can also be expressed in terms of the
%packing fraction $\eta=nV$ as
     %F_{excess} &= k_BT(n_3n_0 + n_2n_1)V  \\
                %&= k_BT\left(\left(\eta\right)\left(n\right) + \left(\frac{3\eta}{R}\right)\left(\frac{3\eta}{4\pi R^2}\right)\right)V  \\      
                %&= k_BT\left(\eta^2+3\eta^2\frac{3}{4\pi R^3}V\right)  \\
                %&= k_BT4\eta^2

@@

     F_{excess} = k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}

@@

where $\vec r_i$ is a position vector of one sphere, $\vec r_j$ is a 
position vector to another sphere, and $|\vec r_i - \vec r_j|$ 
is the distance between the centers of the two spheres.

@@

a convolution of the weight functions which, for a fluid composed of 
spheres of radius R, is given by

@@

    f(|\vec r_i-\vec r_j|)=-\Theta(|\vec r_i-\vec r_j|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})

@@

%where the weights are the same ones given in Equations~\ref{scalor-weight3} through~\ref{scalor-weight0}, 
where the two vector weights are given by

@@

      \vec {w}_2 &= w_2~\frac{\vec r_i-\vec r_j}{|\vec r_i-\vec r_j|} \\
      \vec {w}_1 &= w_1~\frac{\vec r_i-\vec r_j}{|\vec r_i-\vec r_j|}

@@

which are added for the relation to be satisfied, 
and the other weights take the form 
      w_3 &= \Theta(|\vec r_i-\vec r_j|-R)  \label{scalor-weight3} \\
      w_2 &= \delta(|\vec r_i-\vec r_j|-R)  \\
      w_1 &= \frac{w_2}{4\pi |\vec r_i-\vec r_j|}  \\
      w_0 &= \frac{w_2}{4\pi |\vec r_i-\vec r_j|^2} \label{scalor-weight0}

@@

The convolution of the weight functions can be written as

@@

     w_\alpha\otimes w_\gamma(\vec r_i - \vec r_j) &= \int w_\alpha(\vec r') w_\gamma(\vec r_i -\vec r_j - \vec r') d\vec r'   \\                                                             
                                                   &= \int w_\alpha(\vec u - \vec r_j) w_\gamma(\vec r_i -\vec r_j -(\vec u - \vec r_j))d\vec u     \\                                                               
                                                   &= \int w_\alpha(\vec u - \vec r_j) w_\gamma(\vec r_i - \vec u) d\vec u  \\  
                                                   %&= \int w_\gamma(\vec r_a - \vec u)w_\alpha(\vec u - \vec r_b) d\vec u  \\                                                              
                                                   &= \int w_\gamma(-(\vec u - \vec r_i))w_\alpha(\vec u - \vec r_j))  d\vec u  

@@

For the scalor weights this is the same as

@@

                                                   &= \int w_\gamma(\vec u - \vec r_i)w_\alpha(\vec u - \vec r_j)  d\vec u  

@@

and for the vector weights this becomes

@@

                                                   &= -\int \vec {w}_\gamma(\vec u - \vec r_i)\cdot\vec {w}_\alpha(\vec u - \vec r_j)  d\vec u  

@@

%which can also be written as
     %\color{white}w_\alpha\otimes w_\gamma(\vec r_a - \vec r_b)\color{blue} &=\int w_\alpha(\vec x - \vec r_a) w_\gamma(\vec x - \vec r_b) d\vec x
%since the weight functions involve the distances $|\vec r_a - \vec x|=|\vec x - \vec r_a|$.
The lowest order term of the excess Helmholtz free energy can then be written as
     F_{excess} &= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\left(\frac{-f(|\vec r_i-\vec r_j|)}{2}\right)  \\
                &= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}  \\
                &= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~(w_3\otimes w_0+ w_2 \otimes w_1+\vec {w}_2\otimes \vec {w}_1)  \\
                &= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)\left(\int w_0(\vec u - \vec r_i)w_3(\vec u - \vec r_j) d\vec u \color{white}\right)  \nonumber \\
                &+ \color{white}\left(\color{blue}\int w_1(\vec u - \vec r_i)w_2(\vec u - \vec r_j) d\vec x ~-\int \vec {w}_1(\vec u - \vec r_i)\cdot\vec {w}_2(\vec u - \vec r_j) d\vec u\right)    \\
                &= k_BT\int \left[\int n(\vec r_i) w_0(\vec u - \vec r_i)~d\vec r_i \int n(\vec r_j) w_3(\vec u - \vec r_j)~d\vec r_j \color{white}\right]  \nonumber \\
                &+ \int n(\vec r_i) w_1(\vec u - \vec r_i)~d\vec r_i \int n(\vec r_j) w_2(\vec u - \vec r_j)~d\vec r_j  \nonumber \\
                &+ \color{white}\left[\color{blue} \int n(\vec r_i) \vec {w}_1(\vec u - \vec r_i)~d\vec r_i \cdot \int n(\vec r_j) \vec {w}_2(\vec u - \vec r_j)~d\vec r_j \right]\color{blue}d\vec u  \\
                &= k_BT\int n_0n_3 + n_1n_2 - \vec n_1\cdot\vec n_2~d\vec u  \label{Fexlowden-inhomo}\\
                &= k_BT\int \Phi_1 + \Phi_2 + \Phi_3~d\vec u  \\
                &= k_BT\int \Phi(\{n_\alpha(\vec u)\})~d\vec u  \\
                &\text{or replacing u notation with r} \nonumber \\
                &= k_BT\int \Phi(\{n_\alpha(\vec r)\})~d\vec r             
where the weighted densities $n_\alpha$ have the form
    n_\alpha(\vec r)=\int n(\vec r')w_\alpha(\vec r-\vec r')~d\vec r'  
    %n_\alpha(\vec u)=\int n(\vec r')w_\alpha(\vec u-\vec r')~d\vec r'  
For the homogenesous case $\vec n_1\cdot\vec n_2$=0, and Equation~\ref{Fexlowden-inhomo} reduces to Equation~\ref{Fexlowden-homo}.

@@

}\textcolor{blue}{
Lifting the restriction to the low density limit, Rosenfeld derived an
expression for the excess Helmholtz free energy by noting that the higher
order terms in the virial expansion can all be written in terms of an 
"exclusion volume" which is given by Equation~\ref{exclusion-volume}, 
and an "overlap volume".
The exclusion volume is the spherical volume about the center of one sphere 
which the center of another sphere can not enter, and is used to describe
terms with $f_{ij}=f(|\vec r_i - \vec r_j|)$ as has been seen. The overlap volume is 
the volume of intersection between two spheres that overlap, and is used in conjunction 
with the exclusion volume to describe higher order terms in the expansion~\cite{rosenfeld1988}. 
The resulting expression for the excess Helmholtz free energy is given in 
Equation~\ref{FexfromSPT} with $\Phi$'s defined in Equations~\ref{Phi1equation} 
through~\ref{oldPhi3equation}, and
and weighted densities defined in Equation~\ref{weighted-density}, 
and reduces to the same result obtained from Scaled Particle Theory for a 
homogeneous hard-sphere fluid. 

@@

%In order to apply classical Density Functional Theory to a system that has
%a fixed temperature, volume, and number of particles, a functional for 
%the Helmholtz free energy must first be formed. An expression for the ideal 
%portion of the Helmholtz free energy is given by Equation~\ref{eq:Fideal-n(r)}, 
%but forming an expression for the excess Helmholtz free
%energy from Equation~\ref{eq:Fexcess-simplified} for the particular fluid 
%being modeled can be challenging. 
%Soft Fundamental Measure Theory (SFMT) is a method for forming an 
%expression for the excess Helmholtz free energy of a soft-sphere fluid
%that does not depend on explicitly taking into account the interaction 
%that each sphere with every other sphere in the fluid (as required by 
%Equation~\ref{eq:Fexcess-simplified}). In the next chapter on methods, we
%will use an SFMT functional that we adapt to fit the soft-sphere WCA fluid.
%SFMT was developed by Schmidt~\cite{schmidt1999density, schmidt2000fluid} 
%for soft-spheres. It is an extension of Fundamental Measure Theory (FMT) 
%originally developed by 
%Rosenfeld~\cite{rosenfeld1989, rosenfeld1994, rosenfeld1996, fmtfromsptandsignerror} 
%for the hard-sphere fluid, and which expanded 
%Scaled Particle Theory (SPT)~\cite{santos2012phi3} 
%to inhomogeneous, or non-uniform hard-sphere fluids.

Scaled Particle Theory
seeks a general expression for the excess chemical 
potential of a homogeneous hard-sphere fluid which can then be integrated 
to find the excess Helmholtz free energy according to the relation
\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~}\rightarrow{~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{equation}
It is an extension of the work done by Widom who showed that the excess 
chemical potential of a homogeneous fluid corresponds to the reversible 
work of inserting a sphere of radius R into a fluid consisting of hard 
spheres of radius R. 
\begin{align}
	\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}
\end{align}
Widom's result was obtained by considering the difference 
in the excess Helmholtz free energy when one atom is added to an otherwise 
closed system. He showed that the excess chemical potential is related to 
the ensemble average of the Boltzman factor for the potential energy 
difference $V_{diff}$ between a system with N atoms and one with N+1 atoms. 
The resulting equation, shown below, is known as Widom's Insertion 
Formula~\cite{Hansen}.
\begin{align} \label{mu-der-int}
    \mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}
\end{align}
\begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
where 
\begin{align}
    <e^{-\beta{V_{diff}}}>=\sum_s{P_se^{-\beta V_s}}
\end{align}
and
\begin{align}
    Z=\sum_s{e^{-\beta V_s}}
\end{align}

@@

Fundamental Measure Theory can be constructed from Equation~\ref{Fexcess-SPT}  
%and provides an expression for the excess Helmholtz free energy of an 
%inhomogeneous 
%hard-sphere fluid as a functional of the number density profile $n(\vec r)$. 
%
%Soft Fundamental Measure Theory (SFMT), which will be used in this paper, 
%is an extension of FMT and provides an expression for the excess Helmholtz
%free energy of fluids with soft potentials like that of the WCA fluid. 
%It will be helpful to continue looking at FMT before addressing SFMT.
%FMT begins 
by noticing that a quantity $G_{3}$ can be defined such that
\begin{equation}{\eta =\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=nG_{3}}\end{equation}
where $G_{3}$ is the volume of a sphere. 
Likewise, by defining quantities
\begin{equation}{G_{2}=4\pi{R^2}}\end{equation}
\begin{equation}{G_{1}=R}\end{equation}
\begin{equation}{G_{0}=1}\end{equation}
where $G_{2}$ is the surface area of a sphere, and $G_{1}$ is the 
radius of a sphere, Equation~\ref{Fexcess-SPT} can be written as
\begin{equation}\label{FexfromSPTinthermsofG}{F_{excess}=k_{B}T\left(-nG_{0}ln(1-nG_{3})+\frac{nG_{1}nG_{2}}{1-nG_{3}}+\frac{nG_{2}^3}{24\pi(1-Gn_{3})^2}\right)V}\end{equation}
The quantities $G_{3}, G_{2}, G_{1},$ and $G_{0}$ are fundamental 
geometric measures of a sphere, and are what give rise to the name
Fundamental Measure Theory. Introducing a set of variables
%weighted densities $n_{i}=nG_{i}$ 
given by
\begin{equation}\label{n3}{n_{3}=nG_{3}=\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=\eta}\end{equation}
\begin{equation}\label{n2}{n_{2}=nG_{2}=\left(\frac{N}{V}\right)4\pi{R^2}=\frac{3\eta}{R}}\end{equation}
\begin{equation}\label{n1}{n_{1}=nG_{1}=\left(\frac{N}{V}\right)R}\end{equation}
\begin{equation}\label{n0}{n_{0}=nG_{0}=\left(\frac{N}{V}\right)}\end{equation}
Equation~\ref{FexfromSPTinthermsofG} can be written as
\begin{align}\label{FexfromSPT}
    F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}}{1-n_{3}}+\frac{n_{2}^3}{24\pi(1-n_{3})^2}\right)V
\end{align}
which has the form 
\begin{align}
    F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V
\end{align}
%where 
\begin{align}
   \Phi_1 &= -n_{0}\ln(1-n_{3}) \label{Phi1equation} \\
   \Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
   \Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2}   \label{oldPhi3equation}
\end{align}

The quantities $n_{0},n_{1},n_{2},n_{3}$ can be considered as weighted 
densities where the weights are given by $G_{0},G_{1},G_{2},G_{3}$ respectively.
\begin{align}
    n_{i}=nG_{i}
\end{align}
These weights are valid for determining the excess free energy of a system of 
hard spheres with a homogeneous number density n. But these weights, and 
their related  weighted densities, can be written in a more general form 
allowing for their extension to a system with a spatially varying number 
density $n(\vec{r})$, that is, an inhomogeneous fluid, in which case, the
excess Helmholtz free energy is written as
\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation} 
Using a backwards Heaviside step function defined as
\begin{equation}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{equation}
the following expressions are obtained
\begin{equation}{G_{3}=\int_{Volume}{\Theta(|\vec{r}|-R)d{\vec{r}}} = \frac{4}{3}\pi{R^3}}\end{equation}
\begin{equation}{G_{2}=\int_{Volume}{\delta(|\vec{r}|-R)d{\vec{r}}} = 4\pi{R^2}}\end{equation}
\begin{equation}{G_{1}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r}d{\vec{r}}} = R}\end{equation}
\begin{equation}{G_{0}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r^2}d{\vec{r}}} = 1}\end{equation}
where the weighted densities can now be expressed as
\begin{align}
    n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}
\end{align} 
or for a spatially varying number density $n(\vec{r})$,
\begin{equation}\label{weighted-density}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
with weight functions
\begin{equation}\label{eq:w3}{w_{3}=\Theta(|\vec{r}|-R)}\end{equation}
\begin{equation}\label{eq:w2}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}
\begin{equation}\label{eq:w1}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
\begin{equation}\label{eq:w0}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
The weight functions $w_{2}$ and $w_{3}$ have the relation
\begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}
which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.

@@

Tarazona later improved upon the original FMT equations by replacing the
expression for $\Phi_3$ given by Equation~\ref{oldPhi3equation} with the
more accurate expression~\cite{tarazonaphi3, santos2012phi3} 
\begin{align}
\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  \label{eq:Phi3}
\end{align}
in which a tensor weight is incorporated given by
\begin{equation}\label{eq:tensorweight}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
%All the other FMT equations remain the same.
